---
title: "Getting Started with VibeCluster"
author: "VibeCluster Authors"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with VibeCluster}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 7, fig.height = 5,
  out.width = "100%"
)
```

## What is VibeCluster?

**VibeCluster** answers a simple question: *"Given the way a song sounds, what mood does it create — and what other songs will make me feel the same way?"*

It uses three statistical methods from course:

| Method | Role |
|---|---|
| **K-Means Clustering** | Groups songs into *k* mood clusters |
| **PCA** | Projects 6-D features to 2-D for the Mood Map |
| **kNN** | Finds the *n* songs closest to your mood profile |

---

## Step 1 — Load the Dataset

```{r load}
library(VibeCluster)

songs <- load_songs()
print(songs)
```

The built-in dataset has **364 songs** across 6 mood archetypes with 6 audio
features each (energy, valence, danceability, tempo, acousticness, loudness).

---

## Step 2 — Choose *k* with Elbow + Silhouette

```{r summary, fig.height=4.5}
df <- mood_summary(songs, k_range = 2:8)
df
```

Both plots suggest **k = 6** — one cluster per mood archetype. The elbow
levels off after 6, and silhouette peaks there too.

---

## Step 3 — Cluster into Mood Groups

```{r cluster}
result <- mood_cluster(songs, k = 6)
print(result)
```

Each song now has a `mood` column with its cluster label. Cluster centres
are automatically matched to mood names by nearest-archetype distance.

```{r songs_preview}
head(result$songs[, c("title", "artist", "mood", "energy", "valence")])
```

---

## Step 4 — The Mood Map (PCA)

```{r moodmap, fig.height=5.5}
plot_mood_map(result)
```

Each dot is a song; diamonds mark cluster centroids; convex hulls outline
each mood's territory. PC1 captures roughly the *energy–acousticness* axis
and PC2 captures *valence*.

---

## Step 5 — Mood Palette + Feature Profiles

```{r palette, fig.height=5}
pal <- get_mood_palette(result)
print(pal[, c("mood", "count", "proportion")])
```

The heatmap on the right shows each cluster's mean feature profile — you
can see Energetic songs are high energy + high tempo + high loudness, while
Calm songs are high acousticness + low tempo.

---

## Step 6 — Personalised Recommendations (kNN)

Define your mood profile as a 6-element feature vector, then ask for the
closest songs:

```{r recs_angry}
# High-energy, dark mood — Angry/Energetic profile
my_profile <- c(energy        = 0.90,
                valence       = 0.15,
                danceability  = 0.55,
                tempo_norm    = 0.85,
                acousticness  = 0.05,
                loudness_norm = 0.92)

recs <- recommend_songs(result, profile = my_profile, n = 8)
```

```{r recs_calm, fig.height=5}
# Acoustic, gentle profile
calm_profile <- c(energy        = 0.18,
                  valence       = 0.65,
                  danceability  = 0.35,
                  tempo_norm    = 0.20,
                  acousticness  = 0.88,
                  loudness_norm = 0.18)

recs_calm <- recommend_songs(result, profile = calm_profile, n = 8,
                             from_mood = "Calm")
```

The Mood Map highlights your profile (★) and the recommended songs (●) so
you can see where they sit in the mood landscape.

---

## Bringing Your Own Data

Export tracks from the Spotify Web API (or use the `spotifyr` package),
normalise the features, and pass the CSV to `load_songs()`:

```r
my_songs <- load_songs("my_spotify_tracks.csv")
result   <- mood_cluster(my_songs, k = 6)
```

Your CSV only needs the six feature columns; `title` and `artist` are
optional but will appear in recommendation output.

---

## Function Reference

| Function | Description |
|---|---|
| `load_songs()` | Load built-in or custom CSV dataset |
| `mood_cluster()` | K-means clustering into mood groups |
| `plot_mood_map()` | PCA Mood Map scatter plot |
| `get_mood_palette()` | Bar chart + feature heatmap |
| `recommend_songs()` | kNN personalised recommendations |
| `mood_summary()` | Elbow + silhouette plot for choosing k |
